---
title: "Customer Shopping Trends"
author: "Ashley"
date: "2024-11-12"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction:

Online shopping has become more popular, people love to shop online since it is faster and convenient. It is easier to compare prices and research products. You can save yourself time and gas and not have to wait in long lines. All it takes is your phone/computer and your card and you can happily shop online. We are going to look at customer's shopping behaviors from data that I found on Kaggle. We will take a look into specific information like on what people shop for or their prefered season. 

The data I am working with is from Kaggle.
*https://www.kaggle.com/datasets/iamsouravbanerjee/customer-shopping-trends-dataset*
The data is made from Sourav Banerjee, who made the data about a year ago, to explore shopping trends and get a better understand of data analysis.

---

### Loading in the data
```{r message = F}
# loading package to clean up data
library(tidyverse)
# loading package to use for visual plots
library(ggplot2)
library(dplyr)

# loading in the data
Shopping_Trends_Data <-  read.csv("shopping_trends_updated.csv")

Online_Sales_Data <- read.csv("Online Sales Data.csv", stringsAsFactors = T)

```

---

### Showing the unfiltered data
```{r message = F}
# getting a glimpse 
glimpse(Shopping_Trends_Data)
glimpse(Online_Sales_Data)
```

---

## Questions that will be answered:

1. What age group has the highest and lowest purchasing behavior?

2. On average, how much do customers spend per season?

3. Which color of clothes do people buy the most per season?

4. Which payment method do customers perfer using?

5. For each season, what is the most item purchased by males/females?

6. What are the top 5 locations where customers tend to buy from?

7. What category do men shop for the most? What about women?

8. When are sales the greatest and when are they the lowest?

---

## Key terms that will be used
- Age
- Previous Purchases
- Purchase Amount
- Season
- Payment Method
- Item Purchased
- Color
- Location
- Gender
- Category

---

## Summaries of Info
- Total customers: `r nrow(Online_Sales_Data)`

- Product Categories: `r levels(Online_Sales_Data$Product.Category)`

- Region of sales: `r levels(Online_Sales_Data$Region)`

- Start and End Date: `r Online_Sales_Data %>%  summarize("Start Date"= min(as.character(Date)), "End Date"= max(as.character(Date)))`

- Age range of customers: `r min(Shopping_Trends_Data$Age)` and `r max(Shopping_Trends_Data$Age)`
- Percentage of Male customers: `r paste0(sum((Shopping_Trends_Data$Gender == "Male")) / length(unique(Shopping_Trends_Data$Customer.ID)) * 100, "%")`
- Percentage of Female customers: `r paste0(sum((Shopping_Trends_Data$Gender == "Female")) / length(unique(Shopping_Trends_Data$Customer.ID)) * 100, "%")`

---

## 1. Age Groups with highest and lowest purchasing behavior
```{r}


ggplot(Shopping_Trends_Data, aes(x = Age, y = Previous.Purchases)) + geom_density2d()

```

---

## 2) Average people spent per season
```{r}
ggplot(Shopping_Trends_Data %>%
  group_by(Season) %>%
  summarise(mean_by_season = mean(Purchase.Amount..USD.))) + geom_col(mapping = aes(x = Season, y = mean_by_season, fill = Season))

# notate value on the side
```

---

## 3) Colors most purchased per season

ggplot(Shopping_Trends_Data %>%
  group_by(Season) %>%
  summarise(Num_of_Colors = first(names(sort(table(Color), decreasing = TRUE)))))


Top_3_Colors <- Shopping_Trends_Data %>%
  group_by(Season) %>%
  summarise(Num_of_Colors = names(sort(table(Color), decreasing = TRUE))) %>% 
  slice_head(n = 3)

Top_3_Colors


ggplot(Shopping_Trends_Data$Top_3_Colors, aes(x = Customer.ID, y = Color, color = Season)) + geom_point() + facet_wrap(~ Season)



```

---

## 4) Most used Payment Methods
```{r}
ggplot(Shopping_Trends_Data) + geom_bar(mapping = aes(x = Payment.Method, fill = Payment.Method))

```

---

## 5) Items most purchased by gender per season
```{r}


```

---

## 6) Top locations that customers are purchasing from

Top_Locations <- Shopping_Trends_Data %>% 
  arrange(Location) %>% 
  summarise(Top_Locations = names(sort(table(Location), decreasing = TRUE))) %>% 
  slice_head(n = 5)
Top_Locations



ggplot(Shopping_Trends_Data, aes(x = Customer.ID, y = count(Top_Locations), shape = Location)) + geom_point() 





---

## 7) Categories shopped for the most by gender
```{r}


```

---

## 8) Sales behavior summary
```{r}
ggplot(data = Shopping_Trends_Data, aes(x = Customer.ID, y = Category, color = Season)) + geom_point() + facet_wrap(~ Season) 

```